Soundness Proofs for Iterative Deepening

### 2015-11-20 pldi-16/proof/convergence.tex
make the convergence figure but comment it out

The contrapositive statement offers more structure for an easier proof\footnote{
For the reader who likes to avoid non-constructive proofs, % \cite{vargomax}, % TODO: uncomment for camready/TR
note that we are using the constructive contrapositive direction:
Theorem~\ref{thm:convergence} is $A \rightarrow B$ and Theorem~\ref{thm:totalverif} is $\neg B \rightarrow \neg A$.}:

@@

It would be straightforward to show that any other data races in $t_{\alpha{}(i+1)}'$, which would be discovered in this interleaving,
could be reordered to after $\ppothers{\mathcal{I}}{\alpha}{i}'$,
eventually transforming $t_{\alpha{}(i+1)}'$ to contain no conflicting memory accesses but $\ppinstr{\mathcal{I}}{\alpha}{i}$ itself.
Unfortunately, $\mathcal{J}$ is not necessarily reachable, as $\ppothers{\mathcal{I}}{\alpha}{i}'$ still includes any data-race PPs from $\ppothers{\mathcal{I}}{\alpha}{i}$,
which were ordered after $\pai$ in $\mathcal{I}$ and hence not covered by the inductive assumption.
Hence, we must perform the same ``coalescing'' for each thread in $\ppothers{\mathcal{I}}{\alpha}{i}'$ that we did for $\alpha$,
which we will abbreviate:
\[
	\mathcal{K} = [..., (\tai \cup t_{\alpha{}(i+1)}', \paj), \coalesce{\ppothers{\mathcal{I}}{\alpha}{i}'} ]
\]
For each other thread $\beta$, $\coalesce{\ppothers{\mathcal{I}}{\alpha}{i}'}$ will contain a single transition,
starting with the first instruction of $\beta$'s corresponding transition in $\ppothers{\mathcal{I}}{\alpha}{i}$,
and ending with the next sync API PP in $\beta$, which we'll call $\pbk$.
Again, the only certainly-executed instruction by $\alpha$ after $\tai$ is $\ppinstr{\mathcal{I}}{\alpha}{i}$.
It is even possible that neither $\ppinstr{\mathcal{I}}{\alpha}{i}$ nor any other instruction by thread $\alpha$ will conflict with any other thread,
until a different data race PP between two other threads is discovered,
as shown in Figure~\ref{fig:threethreads}.

@@

\begin{figure}[t]
	TODO % TODO
	\caption{In this example, even though thread 1's data-race PP must occur first in a buggy interleaving, it cannot be reached until after Iterative Deepening first reaches the later-occurring data-race PP between threads 2 and 3.}
	\label{fig:threethreads}
\end{figure}

>>>>

The contrapositive statement offers more structure for an easier proof\footnote{
For the reader who likes to avoid non-constructive proofs, % \cite{vargomax}, % TODO CAMREADY: uncomment
note that we are using the constructive contrapositive direction:
Theorem~\ref{thm:convergence} is $A \rightarrow B$ and Theorem~\ref{thm:totalverif} is $\neg B \rightarrow \neg A$.}:

@@

It would be straightforward to show that any other data races in $t_{\alpha{}(i+1)}'$, which would be discovered in this interleaving,
could be reordered to after $\ppothers{\mathcal{I}}{\alpha}{i}'$,
eventually transforming $t_{\alpha{}(i+1)}'$ to contain no conflicting memory accesses but $\ppinstr{\mathcal{I}}{\alpha}{i}$ itself.
Unfortunately, $\mathcal{J}$ is not necessarily reachable, as $\ppothers{\mathcal{I}}{\alpha}{i}'$ still includes any data-race PPs from $\ppothers{\mathcal{I}}{\alpha}{i}$,
which were ordered after $\pai$ in $\mathcal{I}$ and hence not covered by the inductive assumption.
Hence, we must perform the same ``coalescing'' for each thread in $\ppothers{\mathcal{I}}{\alpha}{i}'$ that we did for $\alpha$,
which we will abbreviate:
\[
	\mathcal{K} = [..., (\tai \cup t_{\alpha{}(i+1)}', \paj), \coalesce{\ppothers{\mathcal{I}}{\alpha}{i}'} ]
\]
For each other thread $\beta$, $\coalesce{\ppothers{\mathcal{I}}{\alpha}{i}'}$ will contain a single transition,
starting with the first instruction of $\beta$'s corresponding transition in $\ppothers{\mathcal{I}}{\alpha}{i}$,
and ending with the next sync API PP in $\beta$, which we'll call $\pbk$.
Again, the only certainly-executed instruction by $\alpha$ after $\tai$ is $\ppinstr{\mathcal{I}}{\alpha}{i}$.
It is even possible that neither $\ppinstr{\mathcal{I}}{\alpha}{i}$ nor any other instruction by thread $\alpha$ will conflict with any other thread,
until a different data race PP between two other threads is discovered.
% TODO CAMREADY
%as shown in Figure~\ref{fig:threethreads}.

@@

%\begin{figure*}[t]
%	\small
%	\centering
%\begin{tabular}{rlll}
%	& \multicolumn{3}{c}{\texttt{~~int x = 0, y = 0;~~~~~~~~~~~~~~~~~~~~~~~~~~~}} \\
%	& \multicolumn{3}{c}{\texttt{bool run\_racy\_code = false;~~~~~~~~~~~~~~~~}} \\
%	\\
%	& Thread $\alpha$ & Thread $\beta$ & Thread $\gamma$ \\
%& \texttt{run\_racy\_code = true;} \\
%& \texttt{int tmp = x;} \\
%& // preempted ($p_{\alpha{}1}$) \\
%& & & \texttt{if (run\_racy\_code) \{} \\
%& & & \texttt{~~~~int tmp = y;} \\
%& & & \texttt{~~~~}// preempted ($p_{\gamma{}1}$) \\
%& & \texttt{int tmp = y;} \\
%& & // no preemption ($p_{\beta{}1}$) \\
%& & // $\ppinstr{}{\beta}{1}$ \\
%& & \texttt{y = tmp + 1;} \\
%& & & \texttt{~~~~}// $\ppinstr{}{\gamma}{1}$ \\
%& & & \texttt{~~~~y = tmp + 1;} \\
%& & & \texttt{~~~~if (/* y raced */) \{} \\
%& & & \texttt{~~~~~~~~int tmp = x;} \\
%& & & \texttt{~~~~~~~~}// preempted ($p_{\gamma{}2}$) \\
%& // $\ppinstr{}{\alpha}{1}$ \\
%& \texttt{x = tmp + 1;} \\
%& & & \texttt{~~~~~~~~}// $\ppinstr{}{\gamma}{2}$ \\
%& & & \texttt{~~~~~~~~x = tmp + 1;} \\
%& & & \texttt{~~~~~~~~assert(!(/* x raced */));} \\
%& & & \texttt{~~~~\}} \\
%& & & \texttt{\}} \\
%
%
%\end{tabular}
%	\caption{In this example, even though thread 1's data-race PP must occur first in a buggy interleaving, it cannot be reached until after Iterative Deepening first reaches the later-occurring data-race PP between threads 2 and 3.}
%	\label{fig:threethreads}
%\end{figure*}
